---
title: "Graphics with ggplot"
author: "Sofia"
date: "19/6/2019"
output: html_document
---

## libraries and R global environment
library(tidyverse)

Environment: 1)Tools + 2)Global options + 3) RMarkdonw + 4) Unclick 'show inline for all R Markdowns documents

## exercise chapter 5

Data from: https://github.com/OHI-science/data-science-training
then click on: 1)data + 2)OHI_global_data + 3)raw and 4) copy link inside the 'chunk'

```{r read the data}
library(tidyverse)

ohi_data<-read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_global_data.csv")
```

Explore the data: head(), str(), summary(), dim()

## ggplot (only works with dataframes!!)

```{r start ggplot, echo=FALSE}

ggplot(data = ohi_data, aes(x=georegion_one, y=OHI_score)) + 
  geom_point()

ggplot(data = ohi_data, aes(x=georegion_one, y=OHI_score)) + 
  geom_jitter(width = 0.2)

```
aes stands for aesthetics
geom is the geometric object
geom_point : in vertical lines
geom_jitter : in more sparsed lines

## Boxplot
```{r boxplot}

ggplot(data = ohi_data, aes(x=georegion_one))+
  geom_bar()


```

## Histograms
```{r histograms}
ggplot(data = ohi_data, aes(x=HDI))+
  geom_histogram()
```

Multiple geom exploration
```{r multiple geoms}

ohi_summary<- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_scores_georegion_summary.csv")

ggplot(data=ohi_summary, aes(x=georegions, y=OHI_score_average))+
  geom_bar(stat="identity", col="Blue")+
  geom_jitter(data=ohi_data, aes(x=georegion_one, y=OHI_score, width = 0.2, col="Orange"))


# Aesthetics, either to write it on ggplot r geom_point
ggplot(data = ohi_data, aes(y=OHI_score, x = HDI, color=georegion_one)) +
  geom_point()

ggplot(data = ohi_data) +
  geom_point(aes(y = OHI_score, x = HDI, color=georegion_one))  

## Size

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, size = coastal_pop, color=coastal_pop)) + 
  geom_point()

```
If we are adding more layers it's better to add the things on geom_point instead of using ggplot

Adding a third discreate variable
```{r}
ggplot(data = ohi_data, aes(x=OHI_score, y=HDI, color=georegion_one))+
  geom_point()

ggplot(data = ohi_data, aes(x=OHI_score, y=HDI, shape=georegion_one))+
  geom_point()

```

Themes
```{r}
ggplot(data = ohi_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() 

source('https://raw.githubusercontent.com/OHI-Science/ohiprep/master/src/R/scatterTheme.txt')       
# To make big labes for presentations
ggplot(data = ohi_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() + 
   scatterTheme
```

Change labels
```{r}
ggplot(data = ohi_data, aes(x = OHI_score, y = HDI))+
  geom_point()+
  labs(y="OHI score, 2017",
       x="Human Development Index",
       title = "Countries with high human development have more sustainable oceans",
       color = "Georegion") +  # if color doesn't work, use "fill"
     theme_bw()

```

Colors
```{r}
install.packages("RColorBrewer")
library(RColorBrewer)

display.brewer.all()

my_palette<- brewer.pal(n=9, "YlOrRd")

ggplot(data=ohi_data, aes(x=OHI_score, y=OHI_trend,
                          color=HDI))+
  geom_point(size=3)+
  scale_colour_gradientn(colors=my_palette)

ggplot(data=ohi_data, aes(x=OHI_score, y=OHI_trend,
                          color=HDI))+
  geom_point(size=3)+
  scale_colour_gradientn(colors=c("blue", "red"))

ggplot(data=ohi_data, aes(x=OHI_score, y=OHI_trend,
                          color=georegion_one))+
  geom_point(size=3)+
  scale_color_manual(values=my_palette)

```


